Association between trauma severity models and opportunities for improvement:

### 2022-11-08 manuscript.Rmd
adding new data to text

Karolinska University Hospital is located in Solna, Sweden. The trauma centre at Karolinska University Hospital is equivalent to a level 1 trauma centre according to American College of Surgeons standards [@nksstats] and it is the primary trauma centre for the region of Stockholm. Karolinska University Hospital also has agreements with several other regions to provide trauma care for patients with severe trauma and the trauma centre has thereby a catchment population of almost 3 million residents [@nksstats]. 
<!-- Potential things to add about NKS:
1. How many trauma patients per year?
2. How often do conferences happen, or alternatively, roughly how many per year?
3. Which specialties/professions participate in conferences? 
4. Any other info about conferences at NKS that might be relevant like if there's a particular format or it follows particular guidelines etc 
5. Also if there's any stats about trauma patients at NKS in general, like the avg ISS or similar
... although I guess I have the data to calculate this myself-->

The trauma registry at Karolinska University Hospital is part of SweTrau, Sweden's national trauma registry, and thereby follows SweTrau guidelines. All trauma patients meeting the SweTrau inclusion criteria are registered in the trauma registry by research nurses at the trauma centre. Data is collected from the patient's electronic medical records. The trauma registry contains data on patients from 2012 to 2021. <!-- Double check these dates, maybe give exact dates of registration in the study --> 
<!-- Also is there a time limit for when patients are registered in the registry? Like x days/weeks following admission -->

@@

The trauma care quality database is an internal hospital registry and registers the outcome of M&M conferences. Each case reviewed is registered in the trauma care quality register with data about the outcome of the review. The database contains data on patients from 2014 to 2021. <!-- Double check these dates -->

<!-- How exactly is OFI identified, is it according to specific criteria? -->

@@

### PS

@@

RTS, NORMIT and PS12 were calculated according to the original published algorithms [@rtsalgorithm, @normit, @psalgorithm]. TRISS was calculated using the revised 2009 coefficients [@triss2010].

@@

Participants
------------
A total of `r eligibility[1,2]` patients were recorded in the trauma registry during the study period. Of these, `r eligibility[2,2]` were reviewed at an M&M conference and therefore became eligible for study inclusion. `r post.ofi.exclusion` patients were excluded, due to either being of paediatric age (`r eligibility[3,3]`), being dead on arrival at hospital (`r eligibility[4,3]`), or due to incomplete data (`r eligibility[5,3]`). Figure 2 shows the inclusion process for eligible participants.

```{r}
  exclusion_flowchart(eligibility) ## Only works in HTML
```

@@

```{r}
  as_raw_html(incomplete.data.tbl)
```

Descriptive data
----------------
A total of `r eligibility[5,2]` participants were identified as eligible for inclusion. Their demographical and clinical characteristics are described in Table 1. `r numbers['ofi.true']` cases had at least one OFI and `r numbers['ofi.false']` cases had no OFI. The mean age for all patients was `r numbers['mean.age']` and `r numbers['pc.male']`% were male.

For patients with at least one OFI, the mean TRISS score was `r numbers['t.ofi.t']` and the mean NORMIT score was `r numbers['n.ofi.t']`. For patients with no OFI, the mean TRISS score was `r numbers['t.ofi.f']` and the mean NORMIT score was `r numbers['n.ofi.f']`.

@@

The odds ratio for TRISS was `r or.t['or']` (95% CI: `r or.t[2]`-`r or.t[3]`, p-value: `r stats.t[['or.p']]`). The ROC is plotted in Figure 3. The AUC for the ROC was `r stats.t[['auc']]` (95% CI: `r stats.t[['auc.ci']][1]`-`r stats.t[['auc.ci']][3]`). The integrated calibration index for TRISS was `r stats.t[['ici']]`. The accuracy curve is shown in Figure 4. The best accuracy was `r stats.t[['acc.nrs']][1]` and was found at the cutoff `r stats.t[['acc.nrs']][2]`.

```{r}
plot(stats.t[['roc']], main="Fig 3. TRISS Receiver Operating Characteristic Curve.", colorize = TRUE)
```

```{r}
t.plot.acc <- plot(stats.t[['acc']], main="Fig 4. TRISS Accuracy Curve.")
```

### NORMIT 
The odds ratio for NORMIT was `r or.n['or']` (95% CI: `r or.n[2]`-`r or.n[3]`, p-value: `r stats.n[['or.p']]`). The ROC is plotted in Figure 5. The AUC for the ROC was `r stats.n[['auc']]` (95% CI: `r stats.n[['auc.ci']][1]`-`r stats.n[['auc.ci']][3]`). The integrated calibration index for NORMIT was `r stats.n[['ici']]`. The accuracy curve for NORMIT is shown in Figure 5. The best accuracy was `r stats.n[['acc.nrs']][1]` and was found at the cutoff `r stats.n[['acc.nrs']][2]`.

```{r, echo = FALSE}
plot(stats.n[['roc']], main="Fig 5. NORMIT Receiver Operating Characteristic Curve.", colorize = TRUE)
```

```{r, echo = FALSE}
n.plot.acc <- plot(stats.n[['acc']], main="Fig 6. NORMIT Accuracy Curve.")
```

### PS

@@

>>>>

Karolinska University Hospital is located in Solna, Sweden. The trauma centre at Karolinska University Hospital is equivalent to a level 1 trauma centre according to American College of Surgeons standards [@nksstats] and it is the primary trauma centre for the region of Stockholm. Karolinska University Hospital also has agreements with several other regions to provide trauma care for patients with severe trauma and the trauma centre has thereby a catchment population of almost 3 million residents [@nksstats]. 
The trauma registry at Karolinska University Hospital is part of SweTrau, Sweden's national trauma registry, and thereby follows SweTrau guidelines. All trauma patients meeting the SweTrau inclusion criteria are registered in the trauma registry by research nurses at the trauma centre. Data is collected from the patient's electronic medical records. The trauma registry contains data on patients from 2012 to 2021. <!-- Double check these dates, maybe give exact dates of registration in the study -->

@@

The trauma care quality database is an internal hospital registry and registers the outcome of M&M conferences. M&M conferences are held at Karolinska University Hospital ten times per year and are attended by all involved specialties and professions. Each case reviewed is registered in the trauma care quality register with data about the outcome of the review. The database contains data on patients from 2014 to 2021. <!-- Double check these dates -->

@@

#### PS

@@

RTS, NORMIT and PS12 were computed according to the original published algorithms [@rtsalgorithm; @normit; @psalgorithm]. TRISS was calculated using the revised 2009 coefficients [@triss2010].

@@

Participants
------------
A total of `r exclusion[1,2]` patients were recorded in the trauma registry during the study period. Of these, `r exclusion[2,2]` were reviewed at an M&M conference and therefore became eligible for study inclusion. `r exclusion[7,3]` patients were excluded, due to either being of paediatric age (`r exclusion[3,3]`), being dead on arrival at hospital (`r exclusion[4,3]`), or due to incomplete data (`r exclusion[5,3]`). Figure 2 shows the inclusion process for eligible participants.

```{r}
  exclusion_flowchart(exclusion) ## Only works in HTML
```

@@

```{r}
 as_raw_html(table(na.data, type = "incomplete"))
```

Descriptive data
----------------
A total of `r exclusion[6,2]` participants were identified as eligible for inclusion. Their demographical and clinical characteristics are described in table 2.

The median age was `r dd['age']` (IQR: `r dd['age.lo']`-`r dd['age.hi']`). `r dd['male']`% of participants were male. `r dd['blunt']`% of cases were due to blunt force trauma. Of cases where 30-day survival was available, there was a `r dd['mort']`% 30-day mortality rate (`r dd['survival.na']` cases with missing survival data). Of cases where registered data on the highest level of care was available, `r dd['icu']`% had the ICU as their highest level of care (`r dd['care.na']` cases with missing care level data). The median ISS score was `r dd['iss']` (IQR: `r dd['iss.lo']`-`r dd['iss.hi']`).

At least one OFI was found in `r dd['ofi.true']` of cases (`r dd['ofi.pc']`%). In cases where death occurred within 30 days, at least one OFI was found in `r dd['ofi.mort']`% of cases. The median scores for TRISS, NORMIT, and PS were `r dd['t.med']` (IQR: `r dd['t.lo']`-`r dd['t.lo']`), `r dd['n.med']` (IQR: `r dd['n.lo']`-`r dd['n.hi']`), and `r dd['p.med']` (IQR: `r dd['p.lo']`-`r dd['p.hi']`) respectively.

@@

The odds ratio for TRISS was `r t[['or']]` (95% CI: `r t[['or.ci.lo']]`-`r t[['or.ci.hi']]`, p-value: `r t[['or.p']]`). The ROC is plotted in Figure 3. The AUC for the ROC was `r t[['auc']]` (95% CI: `r t[['auc.ci.lo']]`-`r t[['auc.ci.hi']]`). The integrated calibration index for TRISS was `r t[['ici']]` (95% CI: `r t[['ici.ci']][1]`-`r t[['ici.ci']][2]`. The accuracy curve is shown in Figure 4. The best accuracy was `r t[['acc.max']]` (95% CI: `r t[['acc.ci']][1]`-`r t[['acc.ci']][2]`) and was found at the cutoff `r t[['acc.co']]`.

```{r}
plot(t[['roc']], main="Fig 3. TRISS Receiver Operating Characteristic Curve.", colorize = TRUE)
```

```{r}
plot(t[['acc']], main="Fig 4. TRISS Accuracy Curve.")
```

### NORMIT 
The odds ratio for NORMIT was `r n[['or']]` (95% CI: `r n[['or.ci.lo']]`-`r n[['or.ci.hi']]`, p-value: `r n[['or.p']]`). The ROC is plotted in Figure 3. The AUC for the ROC was `r n[['auc']]` (95% CI: `r n[['auc.ci.lo']]`-`r n[['auc.ci.hi']]`). The integrated calibration index for NORMIT was `r n[['ici']]` (95% CI: `r n[['ici.ci']][1]`-`r n[['ici.ci']][2]`. The accuracy curve is shown in Figure 4. The best accuracy was `r n[['acc.max']]` (95% CI: `r n[['acc.ci']][1]`-`r n[['acc.ci']][2]`) and was found at the cutoff `r n[['acc.co']]`.

```{r, echo = FALSE}
plot(n[['roc']], main="Fig 5. NORMIT Receiver Operating Characteristic Curve.", colorize = TRUE)
```

```{r, echo = FALSE}
plot(n[['acc']], main="Fig 6. NORMIT Accuracy Curve.")
```

### PS
The odds ratio for PS was `r p[['or']]` (95% CI: `r p[['or.ci.lo']]`-`r p[['or.ci.hi']]`, p-value: `r p[['or.p']]`). The ROC is plotted in Figure 3. The AUC for the ROC was `r p[['auc']]` (95% CI: `r p[['auc.ci.lo']]`-`r p[['auc.ci.hi']]`). The integrated calibration index for PS was `r p[['ici']]` (95% CI: `r p[['ici.ci']][1]`-`r p[['ici.ci']][2]`. The accuracy curve is shown in Figure 4. The best accuracy was `r p[['acc.max']]` (95% CI: `r p[['acc.ci']][1]`-`r p[['acc.ci']][2]`) and was found at the cutoff `r p[['acc.co']]`.

```{r}
plot(p[['roc']], main="Fig 6. PS Receiver Operating Characteristic Curve.", colorize = TRUE)
```

```{r}
plot(p[['acc']], main="Fig 7. PS Accuracy Curve.")
```

@@

Table 3 shows a comparison of ?? for the three models analysed.

<!-- Make a table here? after I've run the full script -->

```{r}
  table(results, type = "or.plot")
```
